Competing Bandits Overview

### 2018-05-29 overview/overview.tex
commit a bunch of stuff
updated overview

Considered $K = 3$, $Memory = 100$ (unless otherwise noted) \\
\textbf{The Bandit priors that were considered}:
\begin{itemize}
\item Uniform: Draw the mean rewards for the arms from [0.25, 0.75]
\item ``HeavyTail": We took the mean rewards to be randomly drawn from Beta($\alpha=0.6,\beta=0.6$). With this distribution it was likely to have arms that were at the extremes (close to 1 and close to 0) but also some of the arms with intermediate value means.
\item Needle-in-haystack
\begin{enumerate}
\item High - 2 arms with mean 0.50, 1 arm with mean 0.70 (+ 0.20)
\end{enumerate}
\end{itemize}
\textbf{Algorithms considered}:
\begin{enumerate}
\item ThompsonSampling with priors of $Beta(1, 1)$ for every arm.
\item DynamicGreedy with priors of $Beta(1, 1)$ for every arm
\item Bayesian Dynamic $\epsilon$-greedy with priors of $Beta(1, 1)$ for every arm and $\epsilon=0.05$
\end{enumerate}
\textbf{Agent Algorithms considered}:
\begin{enumerate}
\item HardMax
\item HardMaxWithRandom ($\epsilon = 0.05$)
\item SoftMax ($\alpha = 30$)
\end{enumerate}
\textbf{Memory Sizes}
\begin{enumerate}
\item 100
\end{enumerate}
\pagebreak
\textbf{Competing Bandits Simulation Procedure}
\begin{algorithm}
\begin{algorithmic}[1]
\For{Each prior $p$}
\State Generate true distribution from $p$ (except for needle-in-haystack, just use $p$ itself)
\State Generate $T \times K$ realizations for the arms 
	\For{Each agent algorithm $agent alg$}
		\For{Each principal algorithm pair $principalalg1$, $principalalg2$}
			\For{$N$ simulations}

				\State Give the agents $k$ observations from each principal
				\State Give principal 2 $X$ free observations (the agents also get these observations)
				\State Run simulation for T periods
			\EndFor
		\EndFor
	\EndFor
\EndFor
\end{algorithmic}
\end{algorithm}

@@

@@

Some additional preliminary calibration plots. These plots show the \% of simulations over $N=200$ that identified the best arm over time (based on looking at the posterior distributions at a given time): \\
\includegraphics[scale=0.5]{"needle_in_haystack_best_arm_ident"} \\
\includegraphics[scale=0.5]{"heavy_tail_best_arm_ident"} \\
\includegraphics[scale=0.5]{"uniform_best_arm_identification"}

@@

@@

The first conjecture seems to be confirmed in the simulations. Using the calibration from above ($\epsilon = 0.05$ for HMR and $\alpha = 30$ for SoftMax). However, the HardMax results require some more validation. First, we look at the results for low $T$ (1000) but high $N$ (1200). Then we extend the time horizon, but look at lower $N$. One annoying note, it is hard to compare SoftMax across these next two simulations since in the first $\alpha = 10$ and the in the second $\alpha = 30$

\includepdf[pages={-}]{low_time_high_num_sim}

\includepdf[pages={-}]{tournament_results}

The conjecture for HMR and SM seems to be consistent with the data generated from the simulation (the one exception seems to be HeavyTail for BG vs BEG - more on this later). However, what is going on with HardMax appears confusing. To get some more evidence for the original conjecture, we re-run simulations but track market share over time. In most cases, we get evidence of this and the vast majority of market shares over time look as follows:

\includegraphics[scale=0.30]{decreasing_ms_over_time} \\
\includegraphics[scale=0.30]{standard_ms_over_time} \\
\vspace{0.25cm}

However, there are some simulations where we do not observe this. For instance, \\
\includegraphics[scale=0.30]{odd_ms_over_time} \\

I am not sure what happens in these cases, but one conjecture is that these arise from random noise from a low memory size (memory here is 100, as we have been using in our experiment). I've re-run this experiment for a memory size of 1000 and there are fewer cases which have odd switches as shown above. However, would need to look into this more if we think this is worth digging into more. Another thing that might be interesting here is to ask what happens with HardMax if we get rid of any random noise in the reputation score and make it purely dependent on arm selection. In other words, what happens if the reputation score is constructed by using the true means of the selected arms. This of course is not a realistic model but may be useful for thinking about what is going on in this case. The results for this are reported in the appendix.\\

As a comparison, what do the trajectories for the HardMaxWithRandom agent response functions look like? Haven't dug too deeply into the data for these simulations, but here is a sample one (ThompsonSampling vs DEG - plotted market share is for ThompsonSampling) that is similar to many in those that I have looked at: \\
\includegraphics[scale=0.25]{hmr_over_time} \\
\textbf{Puzzles}: Inconsistent with this interpretation of the HardMax results is the result for the Heavy Tail prior. Why does Thompson Sampling do so poorly! Looking at the reputation plots we see that in the early rounds there is very little difference between the algorithms but Thompson Sampling still does better, even if only slightly. Thus, it's kind of confusing as to what is actually going on. Perhaps worth looking more closely at some of the trajectory plots for this prior. As well, Bayesian Greedy seems to beat Bayesian Epsilon Greedy on this prior even when there is randomness in the agent response model which seems puzzling given the preliminary plots. \\

In general, the conjecture that differences in algorithm performance in early rounds is important for determining results in HardMax and differences in algorithm performance for large $T$ is important for determining results in HardMaxWithRandom and SoftMax seems to be roughly validated (but there are some cases we need to think more about what is happening).

\section*{Incumbent Experiment}
Now, we change $X = 200$. In words, we give principal 2 $X$ free observations (the ``incumbent") and this updates her information set and reputation amongst the agents. In this setting what should the incumbent do? What should the entrant do? Does this depend on the agent model? I would conjecture that when we have sufficient randomness in the agent model the same conclusions as before should hold - namely that the entrant, even if getting beat upon entry, should eventually get enough ``free" observations that she can catch up and take more of the market. However, the time horizon, in contrast to the simultaneous entry case, seems like it may be way longer! So far, most of the simulations have simply looked at $t = 1000$, but it may make sense to look at longer time horizons due to the conjecture from above. \\
\vspace{0.25cm}

@@

\includepdf[pages={-}]{free_obs_tournament}

What next? Have re-run this for lower values of $X$ and higher $T$ (3000) and seen qualitatively similar results. Some potential follow-up questions of interest are:
\begin{enumerate}
\item Does reputation or information play a bigger role in the dominance of the incumbent? Can test this by re-initializing either reputation or information of the incumbent when the entrant wins the market (and re-running simulations of that).
\item Even for HardMax it seems that the entrant gets a non-zero market share (the average seems to be between .05 and .25 usually). What do the market share trajectories look like? Are there a lot of simulations where the entrant simply gets zero and some where the entrant seems to get lucky upon entry and has an increasing market share over time?
\item Run for a larger time horizon for the models with agent randomness to see if better algorithms win in the long-run.
\end{enumerate}

>>>>

Considered $K = 3, 10$, $Memory = 100$ (unless otherwise noted) \\
\textbf{The Bandit priors that were considered}:
\begin{itemize}
\item Uniform: Draw the mean rewards for the arms from [0.25, 0.75]
\item ``HeavyTail": We took the mean rewards to be randomly drawn from Beta($\alpha=0.6,\beta=0.6$). With this distribution it was likely to have arms that were at the extremes (close to 1 and close to 0) but also some of the arms with intermediate value means.
\item Needle-in-haystack
\begin{enumerate}
\item High - $K-1$ arms with mean 0.50, 1 arm with mean 0.70 (+ 0.20)
\end{enumerate}
\end{itemize}
\textbf{Algorithms considered}:
\begin{enumerate}
\item ThompsonSampling with priors of $Beta(1, 1)$ for every arm.
\item DynamicGreedy with priors of $Beta(1, 1)$ for every arm
\item Dynamic $\epsilon$-greedy with priors of $Beta(1, 1)$ for every arm and $\epsilon=0.05$
\end{enumerate}
\textbf{Agent Algorithms considered}:
\begin{enumerate}
\item HardMax
\item HardMaxWithRandom ($\epsilon = 0.05$)
\item SoftMax ($\alpha = 30$)
\end{enumerate}
\textbf{Memory Sizes}
\begin{enumerate}
\item 100
\end{enumerate}
\pagebreak
\textbf{Competing Bandits Simulation Procedure}
\begin{algorithm}
\begin{algorithmic}[1]
\For{Each prior $p$}
\State Generate true distribution from $p$ (except for needle-in-haystack, just use $p$ itself)
\State Generate $T \times K$ realizations for the arms 
	\For{Each agent algorithm $agent alg$}
		\For{Each principal algorithm pair $principalalg1$, $principalalg2$}
			\For{$N$ simulations}
				\State Give principal 2 $X$ free observations (the agents also get these observations)				
				\State Give the agents $k$ observations from each principal
				\State Run simulation for T periods
			\EndFor
		\EndFor
	\EndFor
\EndFor
\end{algorithmic}
\end{algorithm}

@@

\includegraphics[scale=0.5]{"../results/preliminary_figures/Reputation Trajectory for Needle In Haystack High 10 arms"} \\
\includegraphics[scale=0.5]{"../results/preliminary_figures/Reputation Trajectory for Uniform 10 arms"} \\
\includegraphics[scale=0.5]{"../results/preliminary_figures/Reputation Trajectory for Heavy Tail 10 arms"}

@@

Some additional preliminary calibration plots. These plots show the \% of simulations over $N=200$ that identified the best arm over time (based on looking at the posterior distributions at a given time): \\
\includegraphics[scale=0.5]{"preliminary_nih_3_arms"} \\
\includegraphics[scale=0.5]{"preliminary_best_arm_heavy_tail_3"}
\includegraphics[scale=0.5]{"uniform_best_arm_identification_3_arms"} \\
\includegraphics[scale=0.5]{preliminary_nih_10_arms} \\
\includegraphics[scale=0.5]{preliminary_best_arm_uniform_10} 
\includegraphics[scale=0.5]{preliminary_best_arm_heavy_tail_10}

@@

However, this is not always the case. For Heavy Tail with $K = 3$ we have that $DG > DEG$. This does not hold however for $K = 10$. This is possibly because learning is ``easy" for DG in $K = 3$ (it identifies the best arm very quickly and in many cases - see preliminary calibration plots). As well, since DEG engages in non-adaptive exploration and Heavy Tail possibly has some very bad arms (i.e. means close to 0), this non-adaptive exploration could have large effects on reputation. Learning appears to be difficult when $K = 10$ and the reputation gap appears to be slightly higher between DEG and DG but it is unclear what leads to the results we see.

@@

This also does not hold. See the puzzles section after the results.

\includepdf[pages={-}]{many_arms_vs_few_arms}

\textbf{Puzzles}: There are two puzzles with respect to HardMax. Why does Needle in a Haystack ($K = 10$) lead to $TS > DG$ and why does Heavy Tail (both $K = 3$ and $K = 10$) lead to $DG, DEG > TS$? Clearly the original conjecture is flawed in some way. What are things that stand out? Under Heavy Tail we see that the difference in algorithm performance without competition is not that large (looking at reputation plots) and that $DG$ identifies the best arm in a large fraction of runs ($\approx 0.8$). Under Needle in a Haystack we see that the difference between algorithm performance is large and that TS identifies the best arm with only a few samples, but what in particular leads it to win under competition with the HardMax response function?

In general, some properties of the instances and the algorithms seem to drive who wins when. We do not have consistency as in the single instance case where $TS > DEG > DG$, but it's unclear under what conditions we have who wins. Should think more about the properties of the instances.

\section*{Incumbent Experiment}
Now, we change $X = 200$. In words, we give principal 2 $X$ free observations (the ``incumbent") and this updates her information set and reputation amongst the agents. In this setting what should the incumbent do? What should the entrant do? Does this depend on the agent model? I would conjecture that when we have sufficient randomness in the agent model the same conclusions as before should hold - namely that the entrant, even if getting beat upon entry, should eventually get enough ``free" observations that she can catch up and take more of the market. However, the time horizon, in contrast to the simultaneous entry case, seems like it may be way longer! So far, most of the simulations have simply looked at $t = 2000$, but it may make sense to look at longer time horizons due to the conjecture from above. \\
\vspace{0.25cm}

@@

In this version of the simulations we additionally have simulations where the information, reputation, or both are artificially erased from the incumbent. The timing of events is as follows:
\begin{enumerate}
\item The incumbent gets $X$ free observations
\item Either information gains are erased (the posterior beliefs of the incumbent are reset to the fake priors) or the reputation gains are erased (the agents' reputation score for the incumbent is reset) or both
\item Both the incumbent and the entrant get $k=5$ free observations
\end{enumerate}

\includepdf[pages={-}]{reputation_vs_information}

What next? Have re-run this for lower values of $X$ and higher $T$ (3000) and seen qualitatively similar results (see Appendix). Some potential follow-up questions of interest are:
\begin{enumerate}
\item Even for HardMax it seems that the entrant gets a non-zero market share (the average seems to be between .05 and .25 usually). What do the market share trajectories look like? Are there a lot of simulations where the entrant simply gets zero and some where the entrant seems to get lucky upon entry and has an increasing market share over time?
\item Run for a larger time horizon for the models with agent randomness to see if better algorithms win in the long-run.
\item What results are surprising?
\end{enumerate}
